Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-10-05 introduction.tex
macro \tstamp

\begin{example}%[Running example]
\label{ex:running}
We illustrate our framework with a very simplified %toy
running example of \emph{music transcription}:
a given input \emph{timeline} of musical events
from an infinite alphabet $\Sigma$,
is parsed into a structured music score.
Input events of $\Sigma$
are pairs $\mu \ts{\tau}$, where $\mu$ is a
MIDI pitch~\cite{Selfridge-Field97beyondMIDI}, %\cite{MIDIfile}
and $\tau \in \mathbb{Q}$ is a timestamp in seconds.
Such inputs typically correspond
to the recording of a live performance, \eg
$I = (69\ts{0.07}),
     (71\ts{0.72}),
     (73\ts{0.91}),
     (74\ts{1.05}),
     (76\ts{1.36}),
     (77\ts{1.71})$. % over interval $[0,2[$,

The output of parsing is a sequence of
timed symbols $\nu\ts{\tau'}$ in an alphabet $\Delta$,
%events (or \emph{notes})
%in Common Western Music Notation (CWMN)~\cite{Gould11Notation}
where $\nu$  represents
an \emph{event} (or \emph{note}),
specified by its \emph{pitch name}
(\eg, $\mbox{\footnotesize A4}$, $\mbox{\footnotesize G5}$, \etc.),
an event \emph{continuation} (symbol `$-$`, see Example~\ref{ex:SWT}),
or a \emph{markup} (opening or closing parenthesis). %(\emph{parenthesis}),
The temporal information $\tau'$ 
is either a time interval, for the opening parentheses
(representing the duration between the parenthesis 
 and the matching closing one), 
or a timestamp, for the other symbols.
The time points in $\tau'$ belong to a rhythmic ``grid'' obtained from recursive divisions:
whole notes ($\musWhole$) split in halves ($\musHalf$), halves
in quarters ($\musQuarter$), eights ($\musEighth$), \etc.
%
For instance, the output score
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
  %\includegraphics[scale=0.20]{pictures/score5.png},
corresponds to a hierarchical structure
that can be linearized as the sequence
$O =$
$\ccall{\mathsf{m}}\ts{[0,1]}$,
$\ccall{2}\ts{[0,1]}$,
$\mbox{\footnotesize A4}\ts{0}$,
$\ccall{2}\ts{[\frac{1}{2},1]}$,
$-\ts{\frac{1}{2}}$,
$\ccall{2}\ts{[\frac{3}{4},1]}$,
$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$,
$\mbox{\footnotesize C$\sharp$5}\ts{\frac{7}{8}}$,
$\creturn{2}\ts{1}$,
$\creturn{2}\ts{1}$,
$\creturn{2}\ts{1}$,
$\ccall{\mathsf{m}}\ts{[1,2]}$,
$\ccall{3}\ts{[1,2]}$,
$\mbox{\footnotesize D5}\ts{1}$,
$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
$\creturn{3}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$.
%\end{enumerate}
The opening markups $\ccall{\mathsf{m}}$ %and $\creturn{\mathsf{m}}$
delimit \emph{measures},
which are time intervals of duration 1 in this example,
while the subsequences of $O$ between markups~$\ccall{d}$ and~$\creturn{d}$,
for some natural number~$d$,
represent a division of the time interval attached to $\ccall{d}$,
of duration $\ell$,
into $d$ sub-intervals of equal duration $\frac{\ell}{d}$.
%
We will show that $O$ is a solution for the
parsing of $I$. Note that several other parsings are possible
like \eg \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}.
SW-parsing associates a weight value
to each solution, and our framework
aims at selecting the best one with respect to this weight.
\endex
\end{example}

>>>>

\begin{example}%[Running example]
\label{ex:running}
We illustrate our framework with a very simplified %toy
running example of \emph{music transcription}:
a given input \emph{timeline} of musical events
from an infinite alphabet $\Sigma$,
is parsed into a structured music score.
Input events of $\Sigma$
are pairs $\mu \tstamp{\tau}$, where $\mu$ is a
MIDI key number~\cite{Selfridge-Field97beyondMIDI}, %\cite{MIDIfile}
and $\tau \in \mathbb{Q}$ is a timestamp in seconds.
Such inputs typically correspond
to the recording of a live performance, \eg
$I = (69\tstamp{0.07}),
     (71\tstamp{0.72}),
     (73\tstamp{0.91}),
     (74\tstamp{1.05}),
     (76\tstamp{1.36}),
     (77\tstamp{1.71})$. % over interval $[0,2[$,

The output of parsing is a sequence of
timed symbols $\nu\tstamp{\tau'}$ in an alphabet $\Delta$,
%events (or \emph{notes})
%in Common Western Music Notation (CWMN)~\cite{Gould11Notation}
where $\nu$  represents
an \emph{event} (or \emph{note}),
specified by its \emph{pitch name}
(\eg, $\mbox{\footnotesize A4}$, $\mbox{\footnotesize G5}$, \etc.),
an event \emph{continuation} (symbol `$-$`, see Example~\ref{ex:SWT}),
or a \emph{markup} (opening or closing parenthesis). %(\emph{parenthesis}),
The temporal information $\tau'$ 
is either a time interval, for the opening parentheses
(representing the duration between the parenthesis 
 and the matching closing one), 
or a timestamp, for the other symbols.
The time points in $\tau'$ belong to a rhythmic ``grid'' obtained from recursive divisions:
whole notes ($\musWhole$) split in halves ($\musHalf$), halves
in quarters ($\musQuarter$), eights ($\musEighth$), \etc.
%
For instance, the output score
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex1.pdf},
  %\includegraphics[scale=0.20]{pictures/score5.png},
corresponds to a hierarchical structure
that can be linearized as the sequence
$O =$
$\ccall{\mathsf{m}}\tstamp{[0,1]}$,
$\ccall{2}\tstamp{[0,1]}$,
$\mbox{\footnotesize A4}\tstamp{0}$,
$\ccall{2}\tstamp{[\frac{1}{2},1]}$,
$-\tstamp{\frac{1}{2}}$,
$\ccall{2}\tstamp{[\frac{3}{4},1]}$,
$\mbox{\footnotesize B4}\tstamp{\frac{3}{4}}$,
$\mbox{\footnotesize C$\sharp$5}\tstamp{\frac{7}{8}}$,
$\creturn{2}\tstamp{1}$,
$\creturn{2}\tstamp{1}$,
$\creturn{2}\tstamp{1}$,
$\ccall{\mathsf{m}}\tstamp{[1,2]}$,
$\ccall{3}\tstamp{[1,2]}$,
$\mbox{\footnotesize D5}\tstamp{1}$,
$\mbox{\footnotesize E5}\tstamp{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\tstamp{\frac{5}{3}}$,
$\creturn{3}\tstamp{2}$,
$\creturn{\mathsf{m}}\tstamp{2}$,
$\creturn{\mathsf{m}}\tstamp{2}$.
%\end{enumerate}
The opening markups $\ccall{\mathsf{m}}$ %and $\creturn{\mathsf{m}}$
delimit \emph{measures},
which are time intervals of duration 1 in this example,
while the subsequences of $O$ between markups~$\ccall{d}$ and~$\creturn{d}$,
for some natural number~$d$,
represent a division of the time interval attached to $\ccall{d}$,
of duration $\ell$,
into $d$ sub-intervals of equal duration $\frac{\ell}{d}$.
%
We will show that $O$ is a solution for the
parsing of $I$. Note that several other parsings are possible
like \eg \includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}.
SW-parsing associates a weight value
to each solution, and our framework
aims at selecting the best one with respect to this weight.
\endex
\end{example}

### 2021-10-05 main.tex
format llncs
macro \tstamp

% Springer llncs style
% https://www.springer.com/gp/computer-science/lncs/conference-proceedings-guidelines
%\input{main-llncs}

@@

% LIPICS v.2021 style
% https://submission.dagstuhl.de/documentation/authors
\input{main-lipics}

@@

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SW{\textsf{sw}\xspace}
\def\swM{\textsf{swM}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\VPA{\textsf{VPA}\xspace}
\def\SVPA{\textsf{sVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phix{{\Phi_\mathsf{x}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\wei{\mathsf{w}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\newcommand{\weie[1]}{\wei_{#1}^\mathsf{e}}
\def\weiei{\weie[\mathsf{i}]}
\def\weiec{\weie[\mathsf{c}]}
\def\weier{\weie[\mathsf{r}]}
\def\weiex{\weie[\mathsf{x}]}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}
\newcommand{\ts}[1]{\mathop{:}{#1}} % timestamp marker
%\newcommand{\ts}[1]{^{#1}} % timestamp marker

@@

\begin{example}\label{ex:SWT}
We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$
(Ex.~\ref{ex:running} and \ref{distance-time}),
%weighted transducer model
with two states $q_0$ and $q_1$, that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of $\Deltai$ events in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
Given  a performed event $\mu$
and the corresponding notated event $\nu$
(\eg MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
the weight  computed by the \SWT is the time distance between both,
as modeled by transitions $\wei_{11}$ below.
%
The continuation symbol $'-'$
(met for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{})
%that represents continuation
is  skipped with no cost (transitions $\wei_{01}$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with a fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
%\philippe{Comprends pas cette phrase}
%\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

@@

\begin{example}\label{ex:nested-word}
In the nested score representation $O \in \Delta^*$ in Ex.~\ref{ex:running},
$\Deltai$ corresponds to timed notes and continuations,
%the set of music events of the form $<\nu, \tau>$
%where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
%rest symbol R.
and $\Deltac$ and $\Deltar$ contain respectively opening and closing parentheses.
%are used to encode the score structure:
%Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
%(resp. $\rangle$)
%and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
%tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
%is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle %\langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
Another example is the other candidate
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
of transcription of~$I$,
linearized into
$O' =$
$\ccall{\mathsf{m}}\ts{[0,1]}$,
$\ccall{2}\ts{[0,1]}$,
$\mbox{\footnotesize A4}\ts{0}$,
$\ccall{2}\ts{[\frac{1}{2},1]}$,
$-\ts{\frac{1}{2}}$,
$\mbox{\footnotesize B4}\ts{\frac{3}{4}}$,
$\creturn{2}\ts{1}$,
$\creturn{2}\ts{1}$,
$\ccall{\mathsf{m}}\ts{[1,2]}$,
$\ccall{3}\ts{[1,2]}$,
$\mbox{\footnotesize `C$\sharp$5`}\ts{1}$,
$\mbox{\footnotesize D5}\ts{1}$,
$\mbox{\footnotesize E5}\ts{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\ts{\frac{5}{3}}$,
$\creturn{3}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$,
$\creturn{\mathsf{m}}\ts{2}$
(see also Fig.~\ref{fig:score-tree}).
The symbol between quotes $\mbox{\footnotesize `C$\sharp$5`}$
represents an \emph{appogiatura}, %(grace-note),
\ie an ornamental note with theoretical duration~0.
\endex
\end{example}

@@

%Let us assume a third component in the symbols of $\Deltac$
%\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
%quel est "l'associate time interval" d'un symbole. Je ne sais
%pas à quoi fait référence "the following call transition". Je
%n'ose pas trop toucher mais je crois qu'il faut revoir.}
%which is the duration of the associated time interval.
%It is used to check the timestamps in the symbols.
%
For instance, the following call transition starts in state
$q_{{i}/{c}}$, meaning that the current interval has number~$i < c$
amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the interval at the stack's top).
It reads a new time-division symbol~$\ccall{d}$:
\(
\weic\bigl(q_{{i}/{c}},
           \ccall{c}\ts{[\tau, \tau+\ell]}, q,
           \ccall{d}\ts{\iota}, q_{{i+1}/{c}},
           q_{{0}/{d}} \bigr) = \alpha_d
\).
The time interval $\iota$ attached to~$\ccall{d}$ 
must have a starting time $\tau + \frac{i\ell}{c}$,
where~$\tau$ and~$\ell$ are respectively
the starting time and duration of the interval attached to
the previous time-division symbol read $\ccall{c}$ (found on the stack).
%plus the total duration of intervals read so far, $i \ell$, and
%the new current interval duration is~$\frac{\ell}{d}$.
%
And it must have a duration $\frac{\ell}{d}$.

\noindent
The weight of the above transition is $\alpha_d$.
We can penalize \eg triplets compared to duplets with $\alpha_2 < \alpha_3$.
%
\noindent
Along with $\ccall{d}$,
the above transition pushes
the state $q_{{i+1}/{c}}$ on the stack,
in order to start the next sub-interval after reading 
a return symbol $\creturn{d}$,
with the transition:
\(
\weir\bigl(q_{{d}/{d}},
           \ccall{d}\ts{[\tau, \tau+\ell]}, q_{{i+1}/{c}},
           \,\creturn{d}\ts{\tau + \ell}, q_{{i+1}/{c}} \bigr) = \one
\).

\noindent
Reading a musical event $\mu$ is done with:
\(
\weii\bigl(q_{{i}/{d}},
           \ccall{d}\ts{[\tau, \tau+\ell]}, q_{{j}/{c}},
           \mu\ts{\tau + \frac{i\ell}{d}}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
\).
% cas des barres de mesure?
%For the case of a new measure symbol,
%$q_{1/1}$ is a special (initial) state.
The transition to start reading the first measure is:
\(
\weiec\bigl(q_{1/1},
           \ccall{\mathsf{m}}\ts{[0, 1]}, q_{1/1},
           q_{1/1} \bigr) = \one
\),
and for a new measure:
\(
\weic\bigl(q_{1/1},
          \ccall{\mathsf{m}}\ts{[\tau-1, \tau]}, q_{1/1},
          \ccall{\mathsf{m}}\ts{[\tau, \tau+1]}, q_{1/1},
          q_{1/1} \bigr) = \one
\).
\endex
\end{example}

>>>>

% Springer llncs style
% https://www.springer.com/gp/computer-science/lncs/conference-proceedings-guidelines
\input{main-llncs}

@@

% LIPICS v.2021 style
% https://submission.dagstuhl.de/documentation/authors
%\input{main-lipics}

@@

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SW{\textsf{sw}\xspace}
\def\swM{\textsf{swM}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\VPA{\textsf{VPA}\xspace}
\def\SVPA{\textsf{sVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phix{{\Phi_\mathsf{x}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\wei{\mathsf{w}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\newcommand{\weie[1]}{\wei_{#1}^\mathsf{e}}
\def\weiei{\weie[\mathsf{i}]}
\def\weiec{\weie[\mathsf{c}]}
\def\weier{\weie[\mathsf{r}]}
\def\weiex{\weie[\mathsf{x}]}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}
\newcommand{\tstamp}[1]{\mathpunct{:}{\scriptstyle #1}} % timestamp marker
%\newcommand{\ts}[1]{^{#1}} % timestamp marker

@@

\begin{example}\label{ex:SWT}
We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$
(Ex.~\ref{ex:running} and \ref{distance-time}),
%weighted transducer model
with two states $q_0$ and $q_1$, that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of $\Deltai$ events in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
Given  a performed event $\mu$
and the corresponding notated event $\nu$
(\eg MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
the weight  computed by the \SWT is the time distance between both,
as modeled by transitions $\wei_{11}$ below.
%
The continuation symbol $'-'$
(met for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{})
%that represents continuation
is  skipped with no cost (transitions $\wei_{01}$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\tstamp{\tau}, \nu\tstamp{\tau'}, q_0) & = &
     \delta(\mu\tstamp{\tau}, \nu\tstamp{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\tstamp{\tau}, \nu\tstamp{\tau'}, q_0) & = &
     \delta(\mu\tstamp{\tau}, \nu\tstamp{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\tstamp{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\tstamp{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\tstamp{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with a fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
%\philippe{Comprends pas cette phrase}
%\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

@@

\begin{example}\label{ex:nested-word}
In the nested score representation $O \in \Delta^*$ in Ex.~\ref{ex:running},
$\Deltai$ corresponds to timed notes and continuations,
%the set of music events of the form $<\nu, \tau>$
%where $\nu$ is either a note (e.g. A4), the continuation symbol $-$, or a
%rest symbol R.
and $\Deltac$ and $\Deltar$ contain respectively opening and closing parentheses.
%are used to encode the score structure:
%Let $\Deltac = \{\langle, \prec\}$ and  $\Deltar = \{\rangle, \succ\}$ where $\langle$
%(resp. $\rangle$)
%and $\prec$ (resp. $\succ$) correspond to start/end of measures, and start/end of
%tuplets. Then  \includegraphics[scale=0.20]{pictures/score5.png}
%is encoded with the nested word $\langle \prec 0, \frac{3}{4}, \frac{7}{8} \succ \rangle %\langle \prec 1, \frac{4}{3}, \frac{5}{3}  \succ \rangle$
Another example is the other candidate
\includegraphics[scale=0.35,trim=0 5mm 0 0]{pictures/ex2.pdf}
of transcription of~$I$,
linearized into
$O' =$
$\ccall{\mathsf{m}}\tstamp{[0,1]}$,
$\ccall{2}\tstamp{[0,1]}$,
$\mbox{\footnotesize A4}\tstamp{0}$,
$\ccall{2}\tstamp{[\frac{1}{2},1]}$,
$-\tstamp{\frac{1}{2}}$,
$\mbox{\footnotesize B4}\tstamp{\frac{3}{4}}$,
$\creturn{2}\tstamp{1}$,
$\creturn{2}\tstamp{1}$,
$\ccall{\mathsf{m}}\tstamp{[1,2]}$,
$\ccall{3}\tstamp{[1,2]}$,
$\mbox{\footnotesize `C$\sharp$5`}\tstamp{1}$,
$\mbox{\footnotesize D5}\tstamp{1}$,
$\mbox{\footnotesize E5}\tstamp{\frac{4}{3}}$,
$\mbox{\footnotesize F5}\tstamp{\frac{5}{3}}$,
$\creturn{3}\tstamp{2}$,
$\creturn{\mathsf{m}}\tstamp{2}$,
$\creturn{\mathsf{m}}\tstamp{2}$
(see also Fig.~\ref{fig:score-tree}).
The symbol between quotes $\mbox{\footnotesize `C$\sharp$5`}$
represents an \emph{appogiatura}, %(grace-note),
\ie an ornamental note with theoretical duration~0.
\endex
\end{example}

@@

%Let us assume a third component in the symbols of $\Deltac$
%\philippe{Je trouve l'exemple un peu confus. Je ne sais pas
%quel est "l'associate time interval" d'un symbole. Je ne sais
%pas à quoi fait référence "the following call transition". Je
%n'ose pas trop toucher mais je crois qu'il faut revoir.}
%which is the duration of the associated time interval.
%It is used to check the timestamps in the symbols.
%
For instance, the following call transition starts in state
$q_{{i}/{c}}$, meaning that the current interval has number~$i < c$
amongst $c$ sub-intervals of duration~$\ell$ (as indicated by the interval at the stack's top).
It reads a new time-division symbol~$\ccall{d}$:
\(
\weic\bigl(q_{{i}/{c}},
           \ccall{c}\tstamp{[\tau, \tau+\ell]}, q,
           \ccall{d}\tstamp{\iota}, q_{{i+1}/{c}},
           q_{{0}/{d}} \bigr) = \alpha_d
\).
The time interval $\iota$ attached to~$\ccall{d}$ 
must have a starting time $\tau + \frac{i\ell}{c}$,
where~$\tau$ and~$\ell$ are respectively
the starting time and duration of the interval attached to
the previous time-division symbol read $\ccall{c}$ (found on the stack).
%plus the total duration of intervals read so far, $i \ell$, and
%the new current interval duration is~$\frac{\ell}{d}$.
%
And it must have a duration $\frac{\ell}{d}$.

\noindent
The weight of the above transition is $\alpha_d$.
We can penalize \eg triplets compared to duplets with $\alpha_2 < \alpha_3$.
%
\noindent
Along with $\ccall{d}$,
the above transition pushes
the state $q_{{i+1}/{c}}$ on the stack,
in order to start the next sub-interval after reading 
a return symbol $\creturn{d}$,
with the transition:
\(
\weir\bigl(q_{{d}/{d}},
           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{{i+1}/{c}},
           \,\creturn{d}\tstamp{\tau + \ell}, q_{{i+1}/{c}} \bigr) = \one
\).

\noindent
Reading a musical event $\mu$ is done with:
\(
\weii\bigl(q_{{i}/{d}},
           \ccall{d}\tstamp{[\tau, \tau+\ell]}, q_{{j}/{c}},
           \mu\tstamp{\tau + \frac{i\ell}{d}}, q_{{i+1}/{d}} \bigr) = \alpha_\mu
\).
% cas des barres de mesure?
%For the case of a new measure symbol,
%$q_{1/1}$ is a special (initial) state.
The transition to start reading the first measure is:
\(
\weiec\bigl(q_{1/1},
           \ccall{\mathsf{m}}\tstamp{[0, 1]}, q_{1/1},
           q_{1/1} \bigr) = \one
\),
and for a new measure:
\(
\weic\bigl(q_{1/1},
          \ccall{\mathsf{m}}\tstamp{[\tau-1, \tau]}, q_{1/1},
          \ccall{\mathsf{m}}\tstamp{[\tau, \tau+1]}, q_{1/1},
          q_{1/1} \bigr) = \one
\).
\endex
\end{example}
